---
title: "ventenata_AMF_analysis"
author: "Claire Tortorelli"
date: "1/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(DHARMa)
library(emmeans)
library(ggpubr)
```

Read in ventenata data with number of hits for each endophyte category: AM arbuscules; AM vesicles; AM hypha; and dark septate endophyte (DS); and hits with no endophyte (zero)

```{r}
df <- read.csv(here("data/veduAMFdf_all.csv"))

```



```{r}
#convert vegtype to a ordered factor
df$vegtype <- factor(df$vegtype, levels = c("ARRI", "ARAR", "SEEP"))
df$plotID <- factor(df$plotID)
#create all AMF column
df$AMF <- df$AMHypha + df$AMVesicle

#crate column for block
df$blockID <- substr(df$plotID, 5,6)

#creat proportion columns for endophyte categories
df$propAMF <- df$AMF / df$totalHits
df$propDS <- df$DS / df$totalHits

#summarize data
df %>% group_by(vegtype) %>% 
    summarise(meanAMF = mean(propAMF*100), sdAMF = sd(propAMF*100),
                meanDS = mean(propDS*100), sdDS = sd(propDS*100),
meanV = mean(AMVesicle))

summary(df)
```

## Data exploration

```{r}
#create hisogram of data

hist(df$propAMF, breaks = 16)
hist(df$AMHypha / df$totalHits, breaks = 16)
hist(df$AMVesicle / df$totalHits, breaks = 16)
hist(df$propDS, breaks = 16)

#plot relationships between endophytes and root length
plot(df$propAMF ~ df$rootL)
plot(df$propDS ~ df$rootL)

#plot relationships between endophytes and veg type
boxplot(df$propAMF ~ df$vegtype)
boxplot(df$propDS ~ df$vegtype)

#plot relationship between endophytes and distance to shrubs
df$shrubDist2 <- df$shrubDist
#replace >500 with 500 for modeling
df$shrubDist2 <- replace(df$shrubDist2, df$shrubDist2 == ">500", 500)
df$shrubDist2 <- as.numeric(df$shrubDist2)
summary(df$shrubDist2)

plot(df$propAMF ~ df$shrubDist2)
plot(df$propDS ~ df$shrubDist2)

#plot endophyte relationship to ventenata biomass (2019) i.e. invaded status and resident biomass (e.g. site productivity)
plot(df$propAMF ~ df$vedu19_g)
plot(df$propDS ~ df$vedu19_g)

plot(df$propAMF ~ df$resident20_g)
plot(df$propDS ~ df$resident20_g)
```

## Model relationships between endophytes and vegetation type:

SEEP = ephemeral wet meadows/ seeps
ARAR = low sage steppe, Artemisia arbuscula
ARRI = scab-flats, Artemisia rigida

```{r}
anova1 <- aov(propAMF ~ vegtype, data = df)
summary(anova1)

library(emmeans)

anova2 <- aov(propAMF ~ vegtype + rootL, data = df)
summary(anova2)
```

using binomial distribution to model counted proportions
```{r}
#model AMF 
#random intercept term for intercept varying among plots and among plots withing blocks
library(glmmTMB)

library(nlme)


fit1 <- glmmTMB(propAMF ~ vegtype + (1|entryID) + (1|blockID/plotID), 
    data = df,
    family = binomial,
    weights = totalHits)

summary(fit1) #summarize model
simulateResiduals(fit1, plot = T) #simulate and plot residuals

#emmeans
emmeans(fit1, pairwise ~ vegtype, type = "response")

#DS model
fit2 <- glmmTMB(propDS ~ vegtype + (1|entryID) + (1|blockID/plotID), 
    data = df,
    family = binomial,
    weights = totalHits)

summary(fit2) #summarize model
simulateResiduals(fit2, plot = T) #simulate and plot residuals
```

```{r}
# ANOVA
library(car)
car::Anova(fit1)
Anova(fit2)

emmeans(fit1, propAMF ~ vegtype)
```


```{r}
#AMF ~ rootL model
fit3 <- glmmTMB(propAMF ~ rootL + (1|entryID) + (1|blockID/plotID), 
    data = df,
    family = binomial,
    weights = totalHits)

summary(fit3) #summarize model
simulateResiduals(fit3, plot = T)

#AMF ~ DS
fit4 <- glmmTMB(propAMF ~ propDS + (1|entryID) + (1|blockID/plotID), #problematic because as a proportion one has to go up as the other goes down? except for the zeros?
    data = df,
    family = binomial,
    weights = totalHits)

summary(fit4) #summarize model
simulateResiduals(fit4, plot = T)


```

Make some pretty boxplots
```{r}
(pamf <- ggplot(data = df, aes(x = vegtype, y = propAMF * 100)) +
    geom_boxplot()+
     theme_bw(base_size = 13)+
     theme(panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) +
    ylab("AMF (% of root colonized)")+ 
    scale_x_discrete(name = "", labels = c("ARRI" = "scab-flat", "ARAR" = "low sage-steppe", "SEEP" = "wet meadow"))
)

(pds <- ggplot(data = df, aes(x = vegtype, y = propDS * 100)) +
    geom_boxplot()+
     theme_bw(base_size = 13)+
     theme(panel.grid.major = element_blank(),
panel.grid.minor = element_blank()) +
    ylab("DSE (% of root colonized)")+ 
    scale_x_discrete(name = "", labels = c("ARRI" = "scab-flat", "ARAR" = "low sage-steppe", "SEEP" = "wet meadow"))
)
```

```{r fig.height=5, fig.width=4}

ggarrange(pamf, pds, 
          #labels = c("(a)", "(b)"),  hjust = 0.015,
          ncol = 1, nrow = 2)

ggsave("AMF_DSEboxplots.png")
```

